\section{Introduction}
\label{sec:introduction}

Avocado (Persea americana) is a high-value crop that plays a significant role in the global agricultural economy. However, its production is constantly threatened by leaf diseases and pest damage such as leaf burn, leaf spot, rust, powdery mildew, and sap-feeding insects (mealybugs), which can severely reduce fruit yield and quality. Early and accurate disease detection is therefore essential for effective crop management and food security.

Traditional disease diagnosis relies on visual inspection by human experts, which is time‑consuming, subjective, and not scalable to large orchards. In recent years, deep learning, particularly convolutional neural networks (CNNs), has emerged as a powerful tool for plant disease classification, often achieving remarkable accuracy on controlled datasets such as PlantVillage. Nevertheless, several critical challenges remain unsolved, preventing the widespread deployment of these models in real‑world agricultural settings.

First, most high‑performing models are computationally heavy, requiring powerful GPUs that are impractical for edge devices like smartphones or Raspberry Pis. Second, collecting and annotating thousands of disease images per class is expensive and labour‑intensive, yet current deep learning models typically demand large amounts of labelled data. Third, even when a model achieves high accuracy, it often fails to communicate its uncertainty, making it unreliable for high‑stakes decisions. Fourth, the lack of explainability reduces trust from end‑users such as farmers and agronomists.

To address these challenges, we present a comprehensive study that integrates four key contributions:

\begin{enumerate}
    \item \textbf{AvocadoLeaf dataset:} A new dataset of 1,861 images covering five disease categories and healthy leaves, captured under diverse real‑world conditions (variable lighting, multiple angles, natural backgrounds). The dataset exhibits moderate class imbalance (max/min ratio 3.96) and fine‑grained difficulty (Silhouette score 0.0067, intra/inter distance ratio 1.59), reflecting the complexity of field‑based diagnosis.
    
    \item \textbf{Lightweight ConvNeXt‑V2 model:} Among 15 benchmarked architectures (Section~\ref{sec:experiments}), ConvNeXt‑V2 achieves the best accuracy-efficiency trade-off, with only 3.4M parameters, 92.14\% test accuracy, and 90.0\% macro F1, while running at 2.30 ms per image on a single GPU.
    
    \item \textbf{Few‑shot learning evaluation:} To assess data efficiency, we compare standard fine‑tuning with Prototypical Networks (ProtoNet) and Relation Networks. ProtoNet achieves 71.7\% accuracy with only 20 labeled images per class, substantially outperforming fine‑tuning (50.0\%) and demonstrating the viability of metric‑based few‑shot learning for fine‑grained plant disease recognition.
    
    \item \textbf{Uncertainty estimation and OOD detection:} We evaluate calibration using Expected Calibration Error (ECE = 0.0382) and out‑of‑distribution (OOD) detection with Monte Carlo Dropout. The model achieves near‑perfect separation on far OOD (random internet images, AUC 0.998) and strong performance on near OOD (leaves of other crops, AUC 0.895), confirming its reliability under realistic conditions.
\end{enumerate}

The remainder of this paper is organised as follows. Section~\ref{sec:related_work} reviews related work in plant disease classification, few‑shot learning, and uncertainty estimation. Section~\ref{sec:dataset_construction} describes the dataset construction process. Section~\ref{sec:dataset_analysis} analyses the dataset statistics, photometric properties, and intrinsic difficulty. Section~\ref{sec:experiments} presents the architecture comparison (including the proposed ConvNeXt‑V2 model), few‑shot learning, uncertainty estimation, and model interpretability results. Section~\ref{sec:discussion} discusses the findings and limitations, and Section~\ref{sec:conclusion} concludes the paper and outlines future work.