\section{Related Work}
\label{sec:related_work}

\subsection{Plant Disease Classification}
Early works on plant disease recognition used traditional machine learning with hand‑crafted features. The introduction of deep learning, especially convolutional neural networks (CNNs), has significantly improved accuracy. The PlantVillage dataset \cite{hughes2015plant} contains more than 50,000 images of healthy and diseased leaves under controlled conditions, and many studies have reported over 99\% accuracy using models such as VGG, ResNet, and Inception \cite{mohanty2016using}. However, these results are often obtained under ideal imaging conditions (uniform background, fixed illumination), which do not reflect the complexity of real‑field scenarios. Several recent works have emphasized the need for field‑collected datasets \cite{barbedo2019plant}, yet such datasets remain scarce, especially for avocado. Our AvocadoLeaf dataset addresses this gap by providing 1,861 images captured under diverse natural conditions, including variable lighting, multiple angles, and cluttered backgrounds.

% \subsection{Lightweight Architectures for Edge Deployment}
% Deploying deep learning models on resource‑constrained edge devices (e.g., smartphones, Raspberry Pi) requires careful trade‑offs between accuracy, model size, and inference speed. MobileNet \cite{howard2017mobilenets} and EfficientNet \cite{tan2019efficientnet} are popular families of lightweight CNNs that use depthwise separable convolutions and neural architecture search. ConvNeXt \cite{liu2022convnet} modernises standard CNNs by incorporating design elements from vision transformers, achieving a favourable accuracy‑efficiency trade‑off. Several studies have applied lightweight models to plant disease classification \cite{atila2021plant}, but most still rely on relatively large models or require fine‑tuning on large datasets. In this work, we benchmark 15 such architectures and identify ConvNeXt‑V2 as offering the best accuracy‑efficiency trade‑off for edge deployment (Section~\ref{sec:experiments}).

\subsection{Few-shot Learning for Plant Disease Recognition}
Collecting large annotated datasets for every possible disease is impractical. Few‑shot learning aims to recognise new classes from only a handful of labelled examples. Popular approaches include metric‑based methods such as Prototypical Networks (ProtoNet) \cite{snell2017prototypical} and Relation Networks \cite{sung2018learning}, as well as optimisation‑based methods like MAML \cite{finn2017model}. In plant pathology, few‑shot learning has been applied to recognise diseases of tomato, rice, and cassava \cite{arguelles2020few, wang2021few}. However, most existing studies use backbones pre‑trained on ImageNet but still require fine‑tuning or meta‑training. To the best of our knowledge, this is the first work to systematically evaluate ProtoNet and Relation Networks with a frozen ConvNeXt backbone on a fine‑grained avocado leaf disease dataset. Our results show that ProtoNet substantially outperforms conventional fine‑tuning, achieving 71.7\% accuracy with only 20 shots per class.

\subsection{Uncertainty Estimation in Deep Learning}
High accuracy alone is insufficient for high‑stakes applications; a reliable model must also be well calibrated and able to detect out‑of‑distribution (OOD) inputs. Expected Calibration Error (ECE) \cite{guo2017calibration} is a standard metric for assessing calibration. Monte Carlo (MC) Dropout \cite{gal2016dropout} provides a simple yet effective way to estimate predictive uncertainty. In plant disease classification, a few studies have examined calibration \cite{wang2022uncertainty} and OOD detection \cite{gonzalez2020identifying}, but these aspects are often overlooked. Our work contributes a thorough uncertainty analysis of a lightweight model for avocado disease diagnosis, including calibration measurement and OOD detection on both far (random internet images) and near (leaves of other crops) datasets, achieving near‑perfect AUC on far OOD and strong AUC (0.895) on near OOD.

% \subsection{Explainable AI for Plant Diagnosis}
% Understanding why a model makes a particular prediction is crucial for building trust with end‑users. Local Interpretable Model-agnostic Explanations (LIME) \cite{ribeiro2016why} is a popular technique for visualising the regions of an input image that most influence the model's decision. In agriculture, XAI has been used to highlight disease symptoms on leaves \cite{sharma2020explainable}, helping farmers verify model outputs. We apply LIME to ConvNeXt‑V2 and show that the model focuses on disease‑relevant areas (e.g., lesions, spots), thereby increasing interpretability and trust.