\section{Discussion}
\label{sec:discussion}

\textbf{Why ConvNeXt‑V2 works well on a fine‑grained, non‑linearly‑separable dataset.} Section~\ref{sec:dataset_analysis} showed that frozen ImageNet features are not linearly separable for AvocadoLeaf (near-zero Silhouette score, held-out linear-probe accuracy of only 65.48\%), reflecting substantial visual overlap between disease categories. The architecture comparison in Section~\ref{sec:experiments} is consistent with this picture: only fine-tuned, non-linear models reach usable accuracy, and among them ConvNeXt‑V2 achieves the best trade-off (92.14\% accuracy, 0.900 macro F1) with the fewest parameters (3.4M) of any competitive architecture, including its predecessor ConvNeXt‑V1 (27.8M). This suggests that, for this task, ConvNeXt‑V2's Global Response Normalisation and design refinements over ConvNeXt‑V1 translate into a more sample- and parameter-efficient decision boundary rather than raw capacity being the deciding factor, since several much larger models (ViT, SwinTransformer) do not clearly outperform it.

\textbf{Few-shot learning narrows but does not close the data-efficiency gap.} The same non-linear-separability finding motivated comparing fine-tuning against metric-based few-shot learning (Section~\ref{subsec:fewshot}). ProtoNet more than doubles fine-tuning's accuracy at every shot level (e.g. 57.4\% vs. 28.6\% at 20 shots), confirming that a task-adapted embedding space is far more data-efficient than directly fine-tuning a classifier head on a handful of examples. However, ProtoNet's absolute accuracy remains well below the $>$90\% achieved with the full training set, so few-shot learning should currently be viewed as a rapid-onboarding aid for new classes or new orchards rather than a replacement for full-scale data collection. RelationNet's near-chance performance (16--19\%, against a 16.7\% random baseline) is a notable negative result: unlike ProtoNet's fixed distance metric, RelationNet must learn its comparison function from the same limited episodes, and with only 1,861 images available for episodic training this learnable module likely lacks enough data to converge to a useful metric, echoing similar findings for RelationNet on other small, fine-grained datasets.

\textbf{The model is reliable at detecting what it doesn't know, but its raw confidence is only moderately calibrated.} The uncertainty analysis (Section~\ref{sec:experiments}) shows two distinct results that carry different practical implications. Out-of-distribution detection is strong for both far OOD (AUC 0.994) and near OOD (AUC 0.985) inputs, meaning the model can reliably flag non-leaf images and, importantly, leaves from other crop species as "not avocado" rather than silently forcing them into one of the six trained classes -- a valuable safety property for a farmer-facing tool. However, the ECE of 0.0625 is above the common 0.05 threshold for good calibration, so the raw softmax confidence should not yet be read as a literal probability of correctness; a lightweight post-hoc calibration step (e.g. temperature scaling) would likely be needed before confidence scores are surfaced directly to end users.

\textbf{Interpretability evidence is now quantitative and full-test-set, but still leaves failure modes uncharacterised.} The LIME explanations in Section~\ref{sec:experiments} localise to symptomatic regions (scorched lesions for \textit{chaymepla}, powdery patches for \textit{moctrang}) rather than background or leaf-margin artefacts, and the deletion-metric evaluation across all 280 test images (mean confidence drop 0.649, 95\% CI [0.618, 0.678], Appendix~\ref{app:lime_faithfulness}) confirms this is not an artefact of a favourably chosen example: hiding the $\sim$20\% of the image LIME attributes each prediction to collapses the model's confidence by nearly two-thirds on average. The drop is nonetheless smaller for misclassified images (0.516) than correct ones (0.660) and is lowest for \textit{repsap} (0.416), whose diffuse, non-contiguous damage pattern is a harder fit for a fixed-size superpixel mask; a systematic error analysis using the confusion matrix (currently future work, Section~\ref{sec:experiments}) would help clarify whether these lower-faithfulness cases are also where the model is least reliable.

\textbf{A language model is a useful explainer of the CNN's decisions, but not a useful substitute decision-maker or automatic error detector.} Section~\ref{subsec:vlm_explain} showed that a small (2B parameter), free, locally-hosted vision-language model can successfully combine the CNN's prediction with LIME's visual attribution to produce grounded, farmer-readable explanations -- this was the original motivation for adding a language model to the system, and it works: the generated text correctly references both the attributed leaf region and the class-appropriate visual symptom in the examples inspected (Figure~\ref{fig:explain_vlm_showcase}). We separately probed two harder uses of a language model that we do \emph{not} report success on (Appendix~\ref{app:llm} gives the full protocol and results for both). First, letting an LLM/VLM \emph{replace} the CNN's decision for its most uncertain predictions -- rather than merely explain them -- backfired: using either a text-only or a vision-language model this way \emph{monotonically reduced} overall test accuracy as more predictions were escalated to it, from the 92.14\% baseline down to 73.6\% (text-only) and 68.6\% (vision) at a 50\% escalation rate, because the language model's own accuracy on the escalated subset (29--50\%) was consistently well below what ConvNeXt‑V2 achieved on the same subset left to decide alone (57--85\%). Second, we tested whether the explainer's own \texttt{supports\_cnn\_diagnosis} judgement -- without ever overriding the CNN -- could serve as a free-standing signal for automatically flagging likely CNN mistakes: it reached a precision of 0.40 but a much lower recall (0.21) than a zero-cost baseline that simply flags the top half of the escalated subset by predictive entropy (precision 0.24, recall 0.79), and every mistake it caught was already caught by that entropy baseline. So, at the model scale available under a no-budget constraint, a language model earns its place in this system specifically as an explanation layer on top of an unchanged CNN + LIME pipeline, not as a decision-maker or as a substitute for the uncertainty estimation the system already has.

\textbf{Limitations.} Several limitations qualify the conclusions above. First, the dataset, while diverse in imaging conditions, was collected in only three provinces of the Central Highlands, and its class sizes range from 165 to 654 images, so both generalisation to other growing regions and performance on the rarest classes (\textit{repsap}, \textit{gisat}) should be treated cautiously. Second, the expert quality-control review (Section~\ref{sec:dataset_construction}) involved a single expert, so the reported agreement reflects consistency with one reviewer rather than genuine inter-expert reliability. Third, the intrinsic difficulty metrics and linear-probe accuracy were computed on a validation set of only 279 images, which inflates the variance of these estimates. Fourth, this study reports GPU inference latency only; the on-device (e.g. Raspberry Pi, mobile) latency and the structured-pruning results needed to fully substantiate edge-deployment claims are not yet available (Section~\ref{sec:experiments}) and are left as future work rather than assumed.
