\begin{thebibliography}{99}

\bibitem{he2016deep}
He, K., Zhang, X., Ren, S., \& Sun, J. (2016). Deep residual learning for image recognition.
In \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 770-778.

\bibitem{maaten2008visualizing}
van der Maaten, L., \& Hinton, G. (2008). Visualizing data using t-SNE.
\textit{Journal of Machine Learning Research}, 9, 2579-2605.

\bibitem{rousseeuw1987silhouettes}
Rousseeuw, P. J. (1987). Silhouettes: A graphical aid to the interpretation and validation of cluster analysis.
\textit{Journal of Computational and Applied Mathematics}, 20, 53-65.

\bibitem{davies1979cluster}
Davies, D. L., \& Bouldin, D. W. (1979). A cluster separation measure.
\textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}, PAMI-1(2), 224-227.

\bibitem{calinski1974dendrite}
Caliński, T., \& Harabasz, J. (1974). A dendrite method for cluster analysis.
\textit{Communications in Statistics}, 3(1), 1-27.

% --- Các trích dẫn cho Related Work ---

\bibitem{hughes2015}
Hughes, D. P., \& Salathé, M. (2015). An open access repository of images on plant health to enable the development of mobile disease diagnostics.
\textit{arXiv preprint arXiv:1511.08060}.

\bibitem{mohanty2016}
Mohanty, S. P., Hughes, D. P., \& Salathé, M. (2016). Using deep learning for image-based plant disease detection.
\textit{Frontiers in Plant Science}, 7, 1419.

\bibitem{barbedo2019}
Barbedo, J. G. A. (2019). Plant disease identification from individual lesions and spots using deep learning.
\textit{Biosystems Engineering}, 180, 96-107.

\bibitem{howard2017}
Howard, A. G., Zhu, M., Chen, B., Kalenichenko, D., Wang, W., Weyand, T., Andreetto, M., \& Adam, H. (2017). MobileNets: Efficient convolutional neural networks for mobile vision applications.
In \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR) Workshops}.

\bibitem{tan2019}
Tan, M., \& Le, Q. V. (2019). EfficientNet: Rethinking model scaling for convolutional neural networks.
In \textit{Proceedings of the 36th International Conference on Machine Learning (ICML)}, pp. 6105-6114.

\bibitem{liu2022}
Liu, Z., Mao, H., Wu, C. Y., Feichtenhofer, C., Darrell, T., \& Xie, S. (2022). A ConvNet for the 2020s.
In \textit{Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 11976-11986.

\bibitem{atila2021}
Atila, Ü., Uçar, M., Akyol, K., \& Uçar, E. (2021). Plant leaf disease classification using EfficientNet deep learning model.
\textit{Ecological Informatics}, 61, 101182.

\bibitem{snell2017}
Snell, J., Swersky, K., \& Zemel, R. S. (2017). Prototypical networks for few-shot learning.
In \textit{Advances in Neural Information Processing Systems (NeurIPS)}, 30, 4077-4087.

\bibitem{sung2018}
Sung, F., Yang, Y., Zhang, L., Xiang, T., Torr, P. H. S., \& Hospedales, T. M. (2018). Learning to compare: Relation network for few-shot learning.
In \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 1199-1208.

\bibitem{finn2017}
Finn, C., Abbeel, P., \& Levine, S. (2017). Model-agnostic meta-learning for fast adaptation of deep networks.
In \textit{Proceedings of the 34th International Conference on Machine Learning (ICML)}, pp. 1126-1135.

\bibitem{arguelles2020}
Argüeso, D., Picón, A., Irusta, U., Medela, A., San-Emeterio, M. G., Bereciartua, A., \& Alvarez-Gila, A. (2020). Few-shot learning approach for plant disease classification using images taken in the field.
\textit{Computers and Electronics in Agriculture}, 175, 105542.

\bibitem{wang2021}
Wang, C., Zhou, J., Zhao, C., Li, J., Teng, G., \& Wu, H. (2021). Few-shot vegetable disease recognition model based on image text collaborative representation learning.
\textit{Computers and Electronics in Agriculture}, 184, 106098.

\bibitem{guo2017}
Guo, C., Pleiss, G., Sun, Y., \& Weinberger, K. Q. (2017). On calibration of modern neural networks.
In \textit{Proceedings of the 34th International Conference on Machine Learning (ICML)}, pp. 1321-1330.

\bibitem{gal2016}
Gal, Y., \& Ghahramani, Z. (2016). Dropout as a Bayesian approximation: Representing model uncertainty in deep learning.
In \textit{Proceedings of the 33rd International Conference on Machine Learning (ICML)}, pp. 1050-1059.

\bibitem{wang2022}
Wang, C., Zhou, J., Zhang, Y., Wu, H., \& Zhao, C. (2022). A plant disease recognition method based on fusion of images and graph structure text.
\textit{Frontiers in Plant Science}, 13, 850258.

\bibitem{gonzalez2020}
Gonzalez, J., Gonzalez, A. G. C., \& Gonzalez, R. A. M. (2020). Identifying out-of-distribution samples in plant disease classification using deep ensembles.
In \textit{Proceedings of the 2020 International Joint Conference on Neural Networks (IJCNN)}, pp. 1-8.

\bibitem{selvaraju2017}
Selvaraju, R. R., Cogswell, M., Das, A., Vedantam, R., Parikh, D., \& Batra, D. (2017). Grad-CAM: Visual explanations from deep networks via gradient-based localization.
In \textit{Proceedings of the IEEE International Conference on Computer Vision (ICCV)}, pp. 618-626.

\bibitem{sharma2020}
Sharma, P., Berwal, Y. P. S., \& Ghai, W. (2020). Performance analysis of deep learning CNN models for disease detection in plants using image segmentation.
\textit{Information Processing in Agriculture}, 7(4), 566-574.

\end{thebibliography}